---
title: "Extract focal NDVI3d cells"
author: "brouwern@gmail.com"
date: "July 5, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

* Load processed rasters of the Pedernales area  
* Identify cells within the raster that correspond the study sites
* Create a time series of values for each year of study
* Also carry out Maximum value compositing (MVC)

## Libraries
```{r}
library(ggplot2)
library(GGally)
library(reshape2)
library(raster)
library(lubridate)
library(plyr)
```



##  Load data

Load rasters
```{r}
load("./data/data_out/NDVI_bi_month_stacked_raster_pedernales_1993_2015.RData")
```


Load spatial points for 
```{r}
load("./data/data_in/pedernales_spatial_points.RData")
```


Look at spatial points data
```{r}
spdf.all@data
```



## Align spatial points to raster

```{r}
p.all <- spTransform(spdf.all,
                     crs(ndvi.list.stacked[[1]]))
```

Check

```{r}
plot(ndvi.list.stacked[[1]][[1]])
plot(p.all, add = TRUE)
```



## Identify cell that a point lies within
gis.stackexchange.com/questions/174925/how-to-find-the-cell-location-index-of-a-raster-using-lat-long-information



Get cell numbers
```{r}
# get cell number
##thorn-scrub site site 
id.scrub <- extract(ndvi.list.stacked[[1]],
                   SpatialPoints(cbind(p.all@data[1,2],
                                       p.all@data[1,1])), 
                   cellnumbers=TRUE)[1]

##broadleaf forest
id.bl <- extract(ndvi.list.stacked[[1]],
                   SpatialPoints(cbind(p.all@data[3,2],
                                       p.all@data[3,1])), 
                   cellnumbers=TRUE)[1]


## Lloyds site (Lloyd et al 2015 PeerJ)
id.Lloyd<- extract(ndvi.list.stacked[[1]],
                   SpatialPoints(cbind(p.all@data[8,2],
                                       p.all@data[8,1])), 
                   cellnumbers=TRUE)[1]



```


Get rows x column location
```{r}
row.col.scrub <- rowColFromCell(ndvi.list.stacked[[1]],
               id.scrub)

row.col.bl <- rowColFromCell(ndvi.list.stacked[[1]],
               id.bl)


row.col.Lloyd <- rowColFromCell(ndvi.list.stacked[[1]],
               id.Lloyd)
```




# Visualize cell locations
Function to visualize selected cells
```{r}
plot.cell <- function(rast, row.col, color = 2){
  plot(extent(rast, 
            row.col[1], row.col[1], 
            row.col[2],  row.col[2]), 
     add=TRUE, col=color, lwd=3)
}
```


Plot selected cells
```{r}
plot.cell(rast = ndvi.list.stacked[[1]],
          row.col = row.col.scrub)
plot.cell(rast = ndvi.list.stacked[[1]],
          row.col = row.col.bl,
          color = 3)
plot.cell(rast = ndvi.list.stacked[[1]],
          row.col = row.col.Lloyd,
          color = 4)
```





# Build time series

Get meta data for each layer
```{r}
#get names of each layr
layer.names <- names(ndvi.list.stacked)

#extrat year month etc
yr <- gsub(".*([21][01-9][01-9][01-9]).*","\\1",layer.names)
mo.no.char <- gsub(".*([_])([1]{0,1}[01-9])([ab])","\\2",layer.names)
mo.ab <- gsub(".*([_])([1]{0,1}[01-9])([ab])","\\3",layer.names)


```


Extract data values from stacked data; "ts" is for "time series"
```{r}
ts.scrub = as.vector(ndvi.list.stacked[row.col.scrub])
ts.bl = as.vector(ndvi.list.stacked[row.col.bl])
ts.Lloyd = as.vector(ndvi.list.stacked[row.col.Lloyd])

```


Put data from selected cells into into dataframes
```{r}
dat.scrub <- data.frame(NDVI = ts.scrub,
                        yr = yr,
                        mo.no.char = mo.no.char,
                        mo.ab = mo.ab,
                        site = "scrub",
                        stringsAsFactors = F,
                        i = 1:length(ts.scrub))

dat.broadleaf <- data.frame(NDVI = ts.bl,
                        yr = yr,
                        mo.no.char = mo.no.char,
                        mo.ab = mo.ab,
                        site = "broadleaf",
                        stringsAsFactors = F,
                        i = 1:length(ts.bl))


dat.Lloyd <- data.frame(NDVI = ts.Lloyd,
                        yr = yr,
                        mo.no.char = mo.no.char,
                        mo.ab = mo.ab,
                        stringsAsFactors = F,
                        site = "Lloyd",
                        i = 1:length(ts.Lloyd))

#stack dataframes
dat <- rbind(dat.scrub,
             dat.broadleaf,
             dat.Lloyd)



```


Set month as numeric
```{r}
dat$mo.no <- as.numeric(dat$mo.no.char)
```




Convert "a" vs "b" month designation to 0.5 for convenience
```{r}
dat$mo.no0.5 <- dat$mo.no + ifelse(dat$mo.ab == "a",0,0.5)

```


Month abbreviations
```{r}
dat$mo.char <- month.abb[dat$mo.no]
```




Plot by month
```{r}
qplot(y = NDVI,
      x = i,
      data = dat,
      color = site,
      geom = "line")
```



# Add NAs for any missing data

If data hasn't been retrieved from web or a file was corrupted, this will add NAs for those years
```{r}
rng <- range(as.numeric(dat$yr))

dat.dummy <- expand.grid(yr = rng[1]:rng[2],
                         mo.no = 1:12,
                         mo.ab = c("a","b"),
                         site = c("scrub","broadleaf","Lloyd"),
                         stringsAsFactors = F)

dat.dummy$mo.no0.5 <- as.numeric(dat.dummy$mo.no) + 
  ifelse(dat.dummy$mo.ab == "a",0,0.5)

dat2 <- merge(dat, dat.dummy,all =  T)
dat2$i.new <- 1:dim(dat2)[1]

```



Look at 2004, which had problems downloading (July 2017)
```{r}
i.2004 <- which(dat2$yr == "2004")

dat2[i.2004,]
```




## Designate data based on whether its the 1st or 2nd image in for a month

22nd to 7th = ~15 days
7th to 22nd = 15 days

```{r}
dat2$day.num <- ifelse(dat2$mo.ab == "a",7,22)
```




## Calculate Julian datas

Concatenate date
```{r}
mo. <- ifelse(as.numeric(dat2$mo.no) > 9, 
              dat2$mo.no ,
                paste(0,dat2$mo.no, sep = ""))


date. <-  paste(dat2$yr,
       mo.,
      dat2$day.num,
      sep = "-")
```



## Calculate Julian w/in year

yday() function is from lubridate package
```{r}
dat2$julian <- yday(date.)
```


## Sort by Julian w/in year

arrange() is in plyr

```{r}
dat3 <- arrange(dat2, yr, julian)
```




## Elapsed time

Calculate date since very 1st timepoint in data
```{r}

i.scrub <- which(dat3$site == "scrub")
i.bl <- which(dat3$site == "broadleaf")
i.lloyd <- which(dat3$site == "Lloyd")


dat3$days.elapsed  <- NA

dat3$days.elapsed[i.scrub] <- cumsum(dat3$julian[i.scrub])
dat3$days.elapsed[i.bl] <- cumsum(dat3$julian[i.bl])
dat3$days.elapsed[i.lloyd] <- cumsum(dat3$julian[i.lloyd])

```


## Plot full time series
```{r}
qplot(y = NDVI,
      x = days.elapsed,
      data = dat3[,],
      color = site,
      geom = "line")
```


```{r}
i.scrub <- which(dat3$site == "scrub")
qplot(y = NDVI,
      x = julian,
      data = dat3[,],
      color = site,
      group = yr,
      geom = "smooth",
      se = FALSE) +
  facet_grid(site ~ ., scale = "free") +
  geom_smooth(se = F)
```




## Make dates

```{r}
dat3$date <- with(dat3, 
                  paste(day.num,
                        mo.no,
                        yr,sep = "/"))
```


## Save bi monthly time series

Save
```{r}
save(dat3, file = "./data/data_out/NDVI_bimonthly_time_series_by_site.RData")
write.csv(dat3, file = "./data/data_out/NDVI_bimonthly_time_series_by_site.csv",
          row.names = F)
```



## Do monthly maximum value compositing (MVC)


### Explore differences w/in month

Cast by 1st vs 2nd half of month
```{r}
NDVI.MVC.mo <- dcast(data = dat3,
                     formula = yr + mo.no + mo.no.char + mo.char + site ~ mo.ab,
                     value.var = "NDVI",
                     fun.aggregate = max)
```

```{r}
names(NDVI.MVC.mo)[grep("^[ab]",names(NDVI.MVC.mo))] <- c("NDVI.wks.a",
                                                          "NDVI.wks.b")
```



Throws some errors b/c of missing data in 2004
```{r}
subset(NDVI.MVC.mo, yr == "2004")
```


Examine differences
```{r}
plot(NDVI.wks.b ~ NDVI.wks.a, data = NDVI.MVC.mo)
```

### Calculate monthly maximum

```{r}
NDVI.MVC.mo$NDVI.MVC.mo <- apply(NDVI.MVC.mo[,c("NDVI.wks.a","NDVI.wks.b")],
                                 1, max)
```


## Define day

Assing to midpooint of month
```{r}
NDVI.MVC.mo$day.no <- 14
```



## Calculate julian w/in year


Formulate date
```{r}
mo. <- ifelse(as.numeric(NDVI.MVC.mo$mo.no) > 9, 
              NDVI.MVC.mo$mo.no ,
                paste(0,NDVI.MVC.mo$mo.no, sep = ""))



NDVI.MVC.mo$date <-  paste(NDVI.MVC.mo$yr,
       mo.,
      NDVI.MVC.mo$day.no,
      sep = "-")

```


Calcualte Julian (yday function is from lubridate)
```{r}
NDVI.MVC.mo$julian <- yday(NDVI.MVC.mo$date)

```


## Sort by Julian w/in year

arrange() is in plyr

```{r}
NDVI.MVC.mo <- arrange(NDVI.MVC.mo, yr, julian)
```


## Save monthly max value composite


```{r}
save(NDVI.MVC.mo, file = "./data/data_out/NDVI_monthly_MVC_time_series_by_site.RData")
write.csv(NDVI.MVC.mo, file = "./data/data_out/NDVI_monthly_MVC_time_series_by_site.csv",
          row.names = F)
```








## Summarize monthly MVC by season


Indices for the seasons
```{r}
NDVI.MVC.mo$mo.no<- as.numeric(NDVI.MVC.mo$mo.no)

i.win <- which(NDVI.MVC.mo$mo.no %in% c(1,2,3))#1st dry season
i.spr <- which(NDVI.MVC.mo$mo.no %in% c(4,5,6))#1st rain season
i.sum <- which(NDVI.MVC.mo$mo.no %in% c(7,8,9))#2nd dry season
i.fal <- which(NDVI.MVC.mo$mo.no %in% c(10,11,12))#2nd rainy season
```


Calculate season summaries
```{r}
form <- formula(yr + site ~ .)

#winter
winter <- dcast(data = NDVI.MVC.mo[i.win,],
      formula = form,
      value.var = "NDVI.MVC.mo",
      fun.aggregate = mean
        )
names(winter)[dim(winter)[2]] <- "mean.NDVI.win"

#spring
spring <- dcast(data = NDVI.MVC.mo[i.spr,],
      formula = form,
      value.var = "NDVI.MVC.mo",
      fun.aggregate = mean
        )
names(spring)[dim(spring)[2]] <- "mean.NDVI.spr"

#summer
summer <- dcast(data = NDVI.MVC.mo[i.sum,],
      formula = form,
      value.var = "NDVI.MVC.mo",
      fun.aggregate = mean
        )
names(summer)[dim(summer)[2]] <- "mean.NDVI.sum"

#fall
fall <- dcast(data = NDVI.MVC.mo[i.fal,],
      formula = form,
      value.var = "NDVI.MVC.mo",
      fun.aggregate = mean
        )
names(fall)[dim(fall)[2]] <- "mean.NDVI.fall"

#all year
year <- dcast(data = NDVI.MVC.mo[,],
      formula = form,
      value.var = "NDVI.MVC.mo",
      fun.aggregate = mean
        )
names(year)[dim(year)[2]] <- "mean.NDVI.year"

```



Merge each season
```{r}
temp <- merge(winter,spring)
temp <- merge(temp, summer)
temp <- merge(temp, fall)
NDVI.MVC.mo.seasonal.summ <- merge(temp, year)
```




Save
```{r}
write.csv(NDVI.MVC.mo.seasonal.summ, file = "./data/data_out/NDVI_MVC_seasonal_summary_time_series_by_site.csv",
          row.names = F)
```

